Optimal (Multiway) Spatial Joins

### 2024-05-19 msj.tex
yufei

\begin{abstract}
    In a {\em spatial join}, we are given a constant number $k \geq 2$ of sets containing axis-parallel rectangles in a 2D space, denoted as $R_1, R_2, ..., R_k$. The objective is to identify and report all $k$-tuples $(r_1, r_2, ..., r_k) \in R_1 \times R_2 \times ... \times R_k$ where the rectangles $r_1, r_2, ..., r_k$ have a non-empty intersection, i.e., $r_1 \cap r_2 \cap ... \cap r_k \neq \emptyset$. This problem holds significant importance in spatial databases and has been extensively studied for over two decades. We explain how to settle the problem in $O(n \log n + \out)$ time using $O(n)$ space --- regardless of the constant $k$ --- where $n = \sum_{i=1}^k |R_i|$ and $\out$ is the result size (i.e., the total number of $k$-tuples reported). The time complexity is asymptotically optimal in the class of comparison-based algorithms, to which our solution belongs. Our result significantly improves the state of the art, which is an algorithm with running time $O(n \log^{\Theta(k)} n + \out)$.
\end{abstract}

@@

In \cite{kcko22}, Khamis et al.\ investigated a type of joins that extend equi-joins in two ways. First, each attribute value in a relation is an interval (rather than a real value); second, each equality predicate in equi-joins is replaced with an ``non-empty intersection'' predicate on the attributes involved. The $k$-SJ problem can be regarded as a join defined next under the framework of \cite{kcko22}. For each $i \in [k]$, define $R_i$ as a relation over two attributes $X$ and $Y$. For each tuple $\bm{u} \in R_i$, its values $\bm{u}(X)$ and $\bm{u}(Y)$ on the two attributes are both intervals (effectively defining a rectangle). The objective is to output all $(\bm{u}_1, \bm{u}_2, ..., \bm{u}_k) \in R_1 \times R_2 \times ... \times R_k$ such that $\cap_{i=1}^k \bm{u}_i(X) \ne \emptyset$ and $\cap_{i=1}^k \bm{u}_i(Y) \ne \emptyset$.

>>>>

\begin{abstract}
    In a {\em spatial join}, we are given a constant number $k \geq 2$ of sets containing axis-parallel rectangles in a 2D space, denoted as $R_1, R_2, ..., R_k$. The objective is to identify and report all $k$-tuples $(r_1, r_2, ..., r_k) \in R_1 \times R_2 \times ... \times R_k$ where the rectangles $r_1, r_2, ..., r_k$ have a non-empty intersection, i.e., $r_1 \cap r_2 \cap ... \cap r_k \neq \emptyset$. This problem holds significant importance in spatial databases and has been extensively studied for over two decades. We explain how to settle the problem in $O(n \log n + \out)$ time using $O(n)$ space --- regardless of the constant $k$ --- where $n = \sum_{i=1}^k |R_i|$ and $\out$ is the result size (i.e., the total number of $k$-tuples reported). The time complexity is asymptotically optimal in the class of comparison-based algorithms, to which our solution belongs. Our result significantly improves the state of the art, which is an algorithm with running time $O(n \log^{2k} n + \out)$.
\end{abstract}

@@

In \cite{kcko22}, Khamis et al.\ investigated a type of joins that extend equi-joins in two ways. First, each attribute value in a relation is an interval (rather than a real value); second, each equality predicate in equi-joins is replaced with an ``non-empty intersection'' predicate on the attributes involved. The $k$-SJ problem can be converted to solving a join defined next under the framework of \cite{kcko22}. For each $i \in [k]$, define $R_i$ as a relation over two attributes $X$ and $Y$. For each tuple $\bm{u} \in R_i$, its values $\bm{u}(X)$ and $\bm{u}(Y)$ on the two attributes are both intervals (effectively defining a rectangle). The objective is to output all $k$-tuples $(\bm{u}_1, \bm{u}_2, ..., \bm{u}_k) \in R_1 \times R_2 \times ... \times R_k$ satisfying $\bigcap_{i=1}^k \bm{u}_i(X) \ne \emptyset$ and $\bigcap_{i=1}^k \bm{u}_i(Y) \ne \emptyset$. Khamis et al.\ \cite{kcko22} developed an algorithm that can process this join in $O(n \log^{2k} n + \out)$ time, where $n = \sum_{i=1}^k |R_i|$. Their algorithm requires $O(n \log n)$ space.
